---
title: "Stats 506, F18, Problem Set 2"
author: "Xun Wang, xunwang@umich.edu"
date: "October 13,2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# 80: --------------------------------------------------------------------------
```

##Question 1
```{r message=FALSE, q1_1}
tib=readr::read_delim("recs2015_usage.csv",delim =',')
tib[1,1]="estimate"
tib[2,1]="standard error"
cv=qnorm(1-{1-.95}/2)
tib=rbind(tib,tib)
tib[3,1]="CI_lower bound";tib[3,2]=tib[1,2]-cv*tib[2,2];
tib[3,3]=tib[1,3]-tib[2,3];tib[3,4]=tib[1,4]-tib[2,4];tib[3,5]=tib[1,5]-tib[2,5]
tib[4,1]="CI_upper bound";tib[4,2]=tib[1,2]+cv*tib[2,2];
tib[4,3]=tib[1,3]+tib[2,3];tib[4,4]=tib[1,4]+tib[2,4];tib[4,5]=tib[1,5]+tib[2,5]
knitr::kable(tib,digits=5,format="pandoc",caption='**Table 1.** *RECS2015 usage*',
             col.names =c('','Electricity/kwh','Natural gas/100 cubic feet',
                          'Propane/gallons','Fuel oil/gallons'),align="c")
```

\pagebreak

##Question 2: Including Plots
###Part a

Using "fdause" command to import XPT files into Stata and then merging two files 
by "seqn", there are 8305 observatiosn left after deleting not matched 
observations.

###Part b
First, we recode "permanent root fragments" as "permanent" and drop individuals 
for whom upper right 2nd bicuspid tooth was not assessed. Then we see "tooth not 
present" and "permanent" as categorical variable 1, and "primary tooth" as 0. 
Using logistic regression to estimate, we get the relationship between age (in 
months) and the probability that an 1 appears. 
```{r q2_1}
stata_1=read.table("model_1.txt",sep='\t')
stata_1=stata_1[c(-1,-2,-(3:8),-19,-20,-21),]
name_coeff=data.frame(name=c("coefficient","standard error","z value","p-value",
                    "confidence interval"))
name_1=rbind(name_coeff,name_coeff)
stata_1=cbind(stata_1[,1],name_1,stata_1[,2])
knitr::kable(stata_1,digits=5,format="pandoc",caption='**Table 2.** *Logistic 
             Regression* Relationship between age and probability of tooth=1',
             col.names=c('variables','','value'),align="c")
```

Because we select"tooth not present" and "permanent tooth" as 1, so get the 
regression equation:
```{r q2_2}
cat("LP=",0.069678,"*ridagemn",-8.35936,'\n')
cat("log(P(lose primary)/P(primary))=LP",'\n')
cat("log(P(primary)/P(lose primary))=-LP",'\n')
cat("P(primary)=exp(-LP)+/(1+exp(-LP))",'\n')
cat("P(primary)=exp(-0.069678*ridagemn+8.35936)/(1+exp(-0.069678*ridagemn+8.35936))")
```

In the following, we use the fitted model in stata to estimate the ages at which 
25, 50, and 75% of individuals who lose their primary upper right 2nd bicuspid. 
Then we round these results to the nearest month. the results in months are:

104,120,136 respectively. 

We could get a range of representative age values (in year) by taking the 
floor of the smallest one and the ceiling of the largest one. That are:

8,9,10,11,12 years old.

```{r q2_3}
curve(exp(-8.35936+0.069678*x)/(1+exp(-8.35936+0.069678*x)),from=0,to=500,
      xlab='age in moths',ylab='probability of losing primary tooth')
abline(v=c(104,120,136))
```

###Part c
First, we add gender variable into the existing model. After fitting, 
we get BIC of this new model is 1542.055, which is bigger than the the BIC of 
the original model(1533.407). So, we do not add gender into the model. 

Then, we want to improve the model by adding race variable. We manipulate these 
data by creating indictors for four kind of races: Mexican American, Non-Hispanic 
White, Non-Hispanic Black and Other Race. We find that White has the largest 
population, so we see white race as reference. After that, we add other three 
races into model one by one and test their BIC. The results are as following:

Mexican American: BIC=1542.285

Other Race: BIC=1541.932

Non-Hispanic Black: BIC=1529.281<1533.407

So, we add black as a categorical variable into the model.

At last, we come to test the poverty income ratio in demographic sheet. 
BIC of the final model equals 1462.895<1529.281. So, the poverty income ratio
should be retained in the final model.

The regression table is as following:
```{r q2_4}
stata_2=read.table("final_model.txt",sep='\t')
stata_2=stata_2[c(-1,-2,-(3:8),-29,-30,-31),]
name_2=rbind(name_1,name_1)
stata_2=cbind(stata_2[,1],name_2,stata_2[,2])
knitr::kable(stata_2,digits=5,format="pandoc",caption='**Table 3.** *Logistic 
             Regression* Regression table of the final model',
             col.names=c('variables','','value'),align="c")
```

###Part d
In Stata, we could get the adjusted predictions and marginal effects simply by
using "margins" command.

First, we should keep in mind that the representative ages determined in part b 
are 8,9,10,11,12 years old, which equals 96,108,120,132,144 months in our model.

Use margins, we could calculate the Adjusted predctions at the mean: at the mean 
of each variables and at each of the representative ages and produnce a table by 
outreg2 command. We import this table as below:
```{r q2_5}
stata_3=read.table("margin_1.txt",sep='\t')
stata_3=stata_3[c(-1,-2,-3,-29,-30,-31),]
name_3=rbind(name_2,name_coeff)
stata_3=cbind(stata_3[,1],name_3,stata_3[,2])
stata_3[,1]=as.character(stata_3[,1])
stata_3[1,1]="8_at";stata_3[6,1]="9_at";stata_3[11,1]="10_at"
stata_3[16,1]="11_at";stata_3[21,1]="12_at"
knitr::kable(stata_3,digits=5,format="pandoc",caption='**Table 4.** *Logistic 
             Regression* Regression table of Adjusted predctions at the mean. 
             At the mean of black and poverty income ratio, at representative 
             age',col.names=c('age/years','','Adjusted predictions'),align="c")
```

Use margins,dydx command, we could calculate the marginal effects at the mean of 
retained black categorical variables and at the same representative ages as part 
b.The table of marginal effects at the mean:
```{r q2_6}
stata_4=read.table("margin_2.txt",sep='\t')
stata_4=stata_4[c(-(1:28),-54,-55,-56),]
name_4=rbind(name_2,name_coeff)
stata_4=cbind(stata_4[,1],name_4,stata_4[,3])
stata_4[,1]=as.character(stata_4[,1])
stata_4[1,1]="8_at";stata_4[6,1]="9_at";stata_4[11,1]="10_at"
stata_4[16,1]="11_at";stata_4[21,1]="12_at"
knitr::kable(stata_4,digits=5,format="pandoc",caption='**Table 5.** *Logistic 
             Regression* Regression table of Marginal effects at the mean. 
             At the mean of poverty income ratio, at 0 and 1 of black,at 
             representative age',col.names=c('age/months','','margins'),
             align="c")
```

Also use margins,dydx command, we could calculate the average marginal effects of 
black categorical variables and at therepresentative ages 96,108,120,132,144 
months.
```{r q2_7}
stata_5=read.table("margin_3.txt",sep='\t')
stata_5=stata_5[c(-(1:28),-54,-55,-56),]
name_5=rbind(name_2,name_coeff)
stata_5=cbind(stata_5[,1],name_5,stata_5[,3])
stata_5[,1]=as.character(stata_5[,1])
stata_5[1,1]="8_at";stata_5[6,1]="9_at";stata_5[11,1]="10_at"
stata_5[16,1]="11_at";stata_5[21,1]="12_at"
knitr::kable(stata_5,digits=5,format="pandoc",caption='**Table 6.** *Logistic 
             Regression* Regression table of Average marginal effects. 
             At 0 and 1 of black,at representative age',
             col.names=c('age/months','','margins'),align="c")
```

###Part e
Finally, we will refit the final model from part c using svy command. First, we 
need to point out how the survey was designed. Using svyset command, we tells 
stata to treat "wtmec2yr" as weights, "sdmvstra" as strata, "sdmvpsu" as primary 
sampling units. And the "vce(linearized)" means survey should has Taylor linearized variance estimation.

Then we get the regression table of the complexed model based on survey 
environment.
```{r q2_8}
stata_6=read.table("svy_final_model.txt",sep='\t')
stata_6=stata_6[c(-1,-2,-(3:8),-29,-30,-31),]
name_6=rbind(name_1,name_1)
stata_6=cbind(stata_6[,1],name_6,stata_6[,2])
knitr::kable(stata_6,digits=5,format="pandoc",caption='**Table 7.** *Logistic 
             Regression* Regression table of the final model with survey command',
             col.names=c('variables','','value'),align="c")
```

The coefficients of these variables do change but they do not change too much.
While, the p-value of poverty income ratio increases and this variable 
changes from significant to non-significant. I think the differences may because 
of the ignorance of survey design at first. The stratified sampling mechanism and 
the weights of the survey change the z-value and coefficients of these variables.

\pagebreak

##Question 3
```{r warning=FALSE, message=FALSE, q3_1}
source("ps2_q3.R")
```

###Part a
Using "read.xport" command to import XPT files into R and then merging two files 
by "SEQN", there are 8305 observatiosn left after deleting not matched 
observations.

###Part b
First, we recode "permanent root fragments" as "permanent" and drop individuals 
for whom upper right 2nd bicuspid tooth was not assessed. Then we see "tooth not 
present" and "permanent" as categorical variable 1, and "primary tooth" as 0. 
Using logistic regression to estimate, we get the relationship between age (in 
months) and the probability that 1 appears. 
```{r q3_2}
r_1=broom::tidy(model_1)
knitr::kable(r_1,digits=5,format="pandoc",caption='**Table 8.** *Logistic 
             Regression* Relationship between age and probability of tooth=1',
             col.names=c('term','estimate','std.error','statistic','p-value'),
             align="c")

```

Because we select"tooth not present" and "permanent tooth" as 1, so get the 
regression equation as the first equation as below.
If we want to know the relationship between age (in months) and the probability 
of "primary",we need to have an extra step.
```{r q3_3}
cat("LP=",model_1$coefficients[2],"*ridagemn",model_1$coefficients[1],'\n')
cat("log(P(lose primary)/P(primary))=LP",'\n')
cat("log(P(primary)/P(lose primary))=-LP",'\n')
LP=c(round(-model_1$coefficients[2],5),"*ridagemn+",
     round(-model_1$coefficients[1],5))
cat("P(primary)=exp(-LP)+/(1+exp(-LP))",'\n')
cat("P(primary)=exp(",LP,")/(1+exp(",LP,"))")
```

In the following, we use the fitted model in R to estimate the ages at which 
25, 50, and 75% of individuals who lose their primary upper right 2nd bicuspid. 
Then we round these results to the nearest month. The results in months are:
```{r q3_4}
age_25
age_50
age_75
```

We could get a range of representative age values (in year) by taking the 
floor of the smallest one and the ceiling of the largest one. That are:
```{r q3_5}
age_floor:age_ceiling
```

Then we show the plot of this relationship and 25,50,75% quantile.
```{r, fig.align='center', q3_6}
curve(exp(model_1$coefficients[1]+model_1$coefficients[2]*x)/(1+                            exp(model_1$coefficients[1]+model_1$coefficients[2]*x)),from=0,to=500,
      xlab='age in moths',ylab='probability of losing primary tooth')
abline(v=c(104,120,136))
```

###Part c
First, we add gender variable into the existing model. After fitting, 
we get BIC of this new model is:
```{r q3_7}
bic_2
```

which is bigger than the the BIC of the original model:
```{r q3_8}
bic_1
```

So, we do not add gender into the model. 

Then, we want to improve the model by adding race variable. We manipulate these 
data by creating indictors for four kind of races: Mexican American, Non-Hispanic 
White, Non-Hispanic Black and Other Race. We find that White has the largest 
population, so we see white race as reference. After that, we add other three 
races into model one by one and test their BIC. The results are as following:
```{r q3_9}
cat("Mexican American: BIC=",bic_3,'\n')
cat("Other Race: BIC=",bic_4,'\n')
cat("Non-Hispanic Black: BIC=",bic_5,"<1533.407")
```

So, we add black as a categorical variable into the model.

At last, we come to test the poverty income ratio in demographic sheet. 
BIC of the final model equals:
```{r q3_10}
bic_6
```
which is smaller than the BIC value of 1529.281.

So, the poverty income ratio should be retained in the final model.

The regression table is as following:
```{r q3_11}
r_2=broom::tidy(model_6)
knitr::kable(r_2,digits=5,format="pandoc",caption='**Table 9.** *Logistic 
             Regression* Regression table of the final model',
             col.names=c('term','estimate','std.error','statistic','p-value'),
             align="c")
```

###Part d
First, we should keep in mind that the representative ages determined in part b 
is 8,9,10,11,12, which equals 96,108,120,132,144 in our model.

Use the predict function, we could calculate the Adjusted predctions at the mean: 
at the mean of each variables and at each of the representative ages and the result
is:
```{r q3_12}
pred=round(pred,5)
name_7=c("8_at","9_at","10_at","11_at","12_at")
r_3=rbind(name_7,pred)
r_3=cbind(c("age/years","Adajusted Predictions"),r_3)
knitr::kable(r_3,digits=5,format="pandoc",caption='**Table 10.** *Logistic 
             Regression* Regression table of Adjusted predctions at the mean. 
             At the mean of black and poverty income ratio, at representative 
             age',row.names=FALSE,align="c")
```

Continuing, we could calculate the marginal effects at the mean of 
retained black categorical variables and at the same representative ages as part 
b.The table of marginal effects at the mean:
```{r q3_13}
margin_atmean=round(margin_atmean,5)
name_8=c("8_at","9_at","10_at","11_at","12_at")
r_4=rbind(name_8,margin_atmean)
r_4=cbind(c("age/years","margins"),r_4)
knitr::kable(r_4,digits=5,format="pandoc",caption='**Table 11.** *Logistic 
             Regression* Regression table of Marginal effects at the mean. 
             At the mean of poverty income ratio, at 0 and 1 of black,at 
             representative age',row.names=FALSE,align="c")
```

Also use predict command, we could calculate the average marginal effects of 
black categorical variables and at the representative ages 8,9,10,11,12 years.
```{r q3_14}
pred_average=round(pred_average,5)
name_9=c("8_at","9_at","10_at","11_at","12_at")
r_5=rbind(name_9,pred_average)
r_5=cbind(c("age/years","margins"),r_5)
knitr::kable(r_5,digits=5,format="pandoc",caption='**Table 12.** *Logistic 
             Regression* Regression table of Average marginal effects. 
             At 0 and 1 of black,at representative age',row.names=FALSE,align="c")
# 80: --------------------------------------------------------------------------
```


