---
title: "Stats 506, F18, Problem Set 4"
author: "Xun Wang, xunwang@umich.edu"
date: "December 5,2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# 80: --------------------------------------------------------------------------
```

##Question 1
```{r q1_1, message=FALSE}
source("ps4_q1.R")
```

```{r q1_2}
all_time_leader=all_time_leader%>%
                  mutate(player=sprintf("%s %s",nameFirst,nameLast))%>%
                  select(player,debut,birthCountry,Hits)
cap='**Table 1.** All-time leader in hits for each birth country.'
knitr::kable(all_time_leader,format='pandoc',caption=cap,align='r',
             col.names=c("Player","Debut","Country of Birth","Hits"))
```

\pagebreak 

##Question 2
###Part a 
Save the data.frame to results_q4a. It is a data.frame with 112 observations 
and 6 variables.

###Part b 
See results of this question in results_q4b.RData.

###Part c
Print data.frame at the end of the script. 
See results in ps4_q2c-X.Rout file (X = 1, 2, 4).

\pagebreak 

##Question 3
###Part a
Read data into SAS from the folder ./data.

###Part b
Use regular expression of SAS to select observations with ¡°MRI¡± in the 
"hcpcs_description" field and "hcpcs_code" starts with a 7.

###Part c
Select two observations with the highest volume, average payment or total 
payment becasue one of the obseravtions has two highest value.

###Part d
Use proc sql to repeat the process above.

###Part e
We export two csv files ps4_q3c.csv and ps4_q3d.csv. In SAS, we use command
"proc compare" and we could see from the proc compare report in SAS that these 
two results match with criterion 0.00001.
