---
title: "Stats 506, F18, Problem Set 1"
author: "Xun Wang, xunwang@umich.edu"
date: "September 30,2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# 80: --------------------------------------------------------------------------
```

##Question 1
  
  Using bash to execute the shell script, the answers are below:
```{bash}
bash ps1_q1.sh
```
\pagebreak

##Question 2: Including Plots
```{r,message=FALSE, include=FALSE, q2}
source("ps1_q2.R")
```

###Question 2 Part a
```{r q2_a}
title_1='**Table 1.** *Airlines names and percentage:* 
airplines responsible for at least 1% flights from Jan. 1 to Oct. 31, 2013.'
cols_1= c('Airlines Names', 'Percentage(%)')
knitr::kable(carrier_name_list, format='pandoc',digits=2, 
             caption=title_1, col.names = cols_1)
```
    
  According to table 1, 11 listed airlines were responsible for at least 1% of 
the flights departing any of the three NYC airports between January 1 and 
October 31, 2013. And the right column is the percent of thier flights.

###Question 2 Part b

  Because the table is too long, we seperate it into three parts: 
the table for 2013 year, 2014 year and the difference between them.
These tables include airline names, flight numbers, percents and limits 
of confidence intervals.
```{r q2_b_1}
title_2='**Table 2.** *Number, percent and confidence interval:* 
flights in the first 10 months of 2013. '
cols_2= c('Airlines','Number 2013','Percent 2013(%)',
          'CI Lower limit 2013(%)','CI Upper limit 2013(%)')
knitr::kable(tib_2013, format='pandoc',digits=2, 
             caption=title_2, col.names = cols_2)

```

```{r q2_b_2}
title_3='**Table 3.** *Number, percent and confidence interval:* 
flights in the first 10 months of 2014.'
cols_3=c('Airlines','Numbers 2014','Percent 2014(%)',
       'CI Lower limit 2014(%)','CI Upper limit 2014(%)')
knitr::kable(tib_2014, format='pandoc',digits=2, 
             caption=title_3, col.names = cols_3)
```

```{r q2_b_3}
title_4='**Table 4.** *number, percent and confidence interval:* 
flights difference in the first 10 months of 2013 and the first 10 months of 2014. 
Sorted by percent of difference.'
cols_4=c('Airlines','Number_diff','Percent_diff(%)',
       'CI Lowerlimit_diff(%)','CI Upperlimit_diff(%)')
knitr::kable(tib_diff, format='pandoc',digits=2, 
             caption=title_4, col.names = cols_4)
```
 
  Limiting tables to the airlines identified in part a, and sorted by percent 
of difference, it could be observed that the largest increase appeared in 
Delta Air Lines Inc.. And Envoy Air had the largest decrease from year 2013 to 
year 2014 except the missing value of company Endeavor Air Inc.. We could 
get the same result by sorting by R functions.
```{r q2_b_4}
in_name
de_name
```

  Then filtering in R, the below table are airlines showed increase in the percent of 
flights but a decrease in the number of flights. Calculating in R, up to 2013 
October there were 279384 times flight in total but 2014 only had 251809 times.
So decrease in the number of total flights made this phenomenon.
```{r q2_b_5}
title_5='**Table 5.** *Number and percent:* 
flights in the first 10 months of 2013 and the first 10 months of 2014. 
Sorted by increase in percent and a decrease in number.'
cols_5= c('Airlines','Numbers13','Percent13(%)',
          'numbers14','Percent14(%)','number_diff',
          'Percent_diff(%)')
knitr::kable(tib_in_de, format='pandoc',digits=2, 
             caption=title_5, col.names = cols_5)
```


###Question 2 Part c

  Among of the three NYC airports, we produce a table showing the percent 
of flights each airline was responsible for in 2013, limiting the airlines 
in part a, including confidence intervals for estimates.
```{r q2_c_1}
title_6='**Table 6.** *Number,airports,percent and confidence interval:* flights in 2013.'
cols_6= c('year','Airports','Airlines','Percent2013(%)',
          'CI Lowerlimit 2013(%)','CI Upperlimit 2013(%)')
knitr::kable(nyc2013,format='pandoc',digits=2,caption=title_6,col.names = cols_6)
```

  Calculating in R, the following results are largest carrier in airport EWR, JFK,LGA, 
respectively.
```{r q2_c_2}
ewr_name_2013
jfk_name_2013
lga_name_2013
```
\pagebreak
  And getting the table for 2014 using the same ways.
```{r q2_c_3}
title_7='**Table 7.** *Number,airports,percent and confidence interval:* flights in 2014.'
cols_7= c('year','Airports','Airlines','Percent2014(%)',
          'CI Lowerlimit 2014(%)','CI Upperlimit 2014(%)')
knitr::kable(nyc2014,format='pandoc',digits=2,caption=title_7,col.names = cols_7)
```

  Similarly, we get the largest carrier in airports EWR, JFK,LGA in 2014, respectively.
```{r q2_c_4}
ewr_name_2014
jfk_name_2014
lga_name_2014
```
\pagebreak

##Question 3: Including Plots
```{r,q3, message=FALSE,include=FALSE}
source("ps1_q3.R")
```

###Question 3 Part a
```{r q3_a_1}
title_8='**Table 8.** *Divisions, percentage and BRR standard error:* stucco as 
the major outside wall material within each division.
Sorted by percent of stucco.'
cols_8= c('Divisions', 'Percentage of stucco(%)', 'standard error','relative standard error')
knitr::kable(recs_decode_1, format='pandoc',digits=2, 
             caption=title_8, col.names = cols_8)
```

  From the table above, we could see percent of homes having stucco construction 
as their major outside wall material within each division. Obviously, Mountain South
had the largest proportion of stucco homes and East South Central had the lowest 
percent. It is the same with extracting their names in R.
```{r q3_a_2}
high_stucco
low_stucco
```


###Question 3 Part b
  
  The table 9 below shows the average total electricity usage in kilowatt hours in 
each division and also show the result stratified by urban and rural status in 
table 10.
```{r q3_b_1}
title_9='**Table 9.** *Divisions, average electricity usage and BRR standard error:* 
in total.'
cols_9= c('Divisions', 'Average usage of electricity', 'standard error','relative standard error')
knitr::kable(recs_ave_total, format='pandoc',digits=2, 
             caption=title_9, col.names = cols_9)
```

```{r q3_b_2}
title_10='**Table 10.** *Divisions, average electricity usage and BRR standard error:* 
stratified by urban and rural status.'
cols_10= c('Divisions', 'status','Average usage of electricity', 'standard error','relative standard error')
knitr::kable(recs_decode_2, format='pandoc',digits=2, 
             caption=title_10, col.names = cols_10)
```


###Question 3 Part c

  In order to get the largest disparity between these proportions,  we calculate the 
proportion of homes with internet access of urban and rural areas first,and the BBR 
standard error to estimate these proportions repectively.

```{r q3_c_1}
title_11='**Table 11.** *Divisions, proportion of internet access and BRR standard 
error.*'
cols_11= c('Divisions', 'status','Proportion of internet access(%)', 'standard error','relative standard error')
knitr::kable(recs_decode_3, format='pandoc',digits=2, 
             caption=title_11, col.names = cols_11)
```

  And then we get their disparity through spreading the tibble.
```{r q3_c_2}
title_12='**Table 12.** *Divisions, disparity of proportion of internet access 
and BRR standard error:* Sorted by disparity.'
cols_12= c('Divisions', 'Proportion of rural status(%)',
           'Proportion of urban status(%)',
           'Disparity of proportion(%)')
knitr::kable(recs_disp, format='pandoc',digits=2, 
             caption=title_12, col.names = cols_12)
```

  The largest disparity happened in:
```{r q3_c_3}
prop_disp
```

